---
title: "Pasture Potential"
runtime: shiny
output: html_document
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
#### load libraries ####
library(tidyverse)
library(leaflet) # interactive map
#library(ggmap) # googlemap
library(cowplot)
library(geosphere)
library(proj4) 
library(KernSmooth) # for contouring
library(sp) # spatial ploygons
library(quantreg)
#library(broom) # used to convert spatial data to tibble
library(rgdal) 

#### these objects are available to all sessions ####
# this is good place to put global data or utility functions
# you can change them (for every session) but you have to use the <<- operator

#### read multiyear data and tidy up ####
file_name <- "DairyBaseForPowerBI.csv"
data_all <- read_csv(file_name, na=c("","NaN","NA"), col_types=cols(Concatbest="c")) # force "Concatbest" to chr
n <- names(data_all)
data_all$Region <- as.factor(data_all$Region)
data_all$Season <- as.factor(data_all$Season)

# check duplicate columns for inconsistencies
# mismatch <- which(data_all$"DBID" != data_all$"DBID_1")
# mismatch <- which(data_all$"Dairy Company(s)" != data_all$"Dairy Company(s)_1") # INCONSISTENT!

# rename useful columns
n[n=="Pasture and Crop eaten t DM/ha"] <- "pasture_eaten"
n[n=="Region"] <- "region"
n[n=="Season"] <- "season"
n[n=="SupplyClean"] <- "supply_number"
names(data_all) <- n

# remove rows missing essential information
data_all <- data_all %>% 
  select(pasture_eaten, region, season, supply_number, long, lat) %>%
  filter((pasture_eaten>0) && (long>0) && (lat<0)) %>%
  drop_na()

# create season list
seasons <- sort(unique(data_all$season))

# create data point map/contours for plotting on leaflet
# https://gis.stackexchange.com/questions/168886/r-how-to-build-heatmap-with-the-leaflet-package
data_pts <- unique(data_all[c("long", "lat")])
kde <- bkde2D(data.matrix(data_pts), bandwidth=c(0.1, 0.1), gridsize=c(100,100))
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat) # contour lines (list)
LEVS <- as.factor(sapply(CL, `[[`, "level")) # contour levels (vector)
NLEV <- length(levels(LEVS)) # number of levels 
pgons <- lapply(1:length(CL), function(i) # convert to polygons (ID=i is actually the level)
    Polygons(list(Polygon(cbind(CL[[i]]$x, CL[[i]]$y))), ID=i))
spgons = SpatialPolygons(pgons)
spgon_cols <- topo.colors(NLEV, NULL)[LEVS]

# convert spatialPolygons to data frame for use in ggplot
#spgonsdf <- tidy(spgons, region=ID)

# calculate NZTM2000 coordinates for farm locations
proj4string <- "+proj=tmerc +lat_0=0.0 +lon_0=173.0 +k=0.9996 +x_0=1600000.0 +y_0=10000000.0 +datum=WGS84 +units=m"
nzgd <- data.matrix(data_all[,c("long", "lat")])
nztm <- project(xy=nzgd, proj=proj4string)
temp <- project(xy=nztm, proj=proj4string, inverse=TRUE)
data_all$east <- nztm[,1]
data_all$north <- nztm[,2]
# relative locations
# data_all$x <- (nztm[,1]-nztm[rr,1])/1000 
# data_all$y <- (nztm[,2]-nztm[rr,2])/1000

#### read soil layer ####
soil <- readOGR(".", "fsl-new-zealand-soil-classification") #from Hans

#### define some constants ####
trim <- 0.0 # rqss fails near tails if insufficient data
probs <- seq(trim, 1-trim, 0.02)
nprobs <- length(probs)
windows <- c(60,40,20)
minn <- 4L # minimum number of farms in a window for analysis
cb9sron <- c('#88CCee', '#CC6677', '#DDCC77', '#117733', '#332288', '#AA4499', '#44AA99', '#999933', '#882255')
nz <- map_data("nz") # nz coastline data

# gets a list of default ggplot colours
gg_colour_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
window_cols <- gg_colour_hue(length(windows))

```

Pasture is a fundamental component of profitable dairy systems. In general, the more pasture you can grow and feed to your animals, the more profitable your system will be. This tool allows you to compare your pasture produced and consumed with pasture produced and consumed on other farms in your region. This can indicate whether there is potential for you to increase pasture production and intake, and hence profitability.

# Where are You?

Please select which season of data to analyse, and click your location on the map. The contours show the availability of data.

```{r echo=FALSE, eval=TRUE}
#### shiny ui widgets to get inputs (reactive values) ####
# Note: Unlike Shiny Apps, Interactive R Markdown Documents are do not require ui and server, 
#       the whole document is treated as a server.
  
  # Note: my itself is not a reactive object!
  my <- reactiveValues(name='You Are Here', 
                       # long=172.833333, lat=-41.5, # Nelson
                       long=175.619105, lat=-40.386396, # Massey
                       # long=174.865530, lat=-41.259256, # Wellington
                       # long=175.352116, lat=-37.781841, # DairyNZ
                       distortion=1, east=NA, north=NA, season=tail(seasons, 1))
  
  v <- reactiveValues(zoom=5, minzoom=5, maxzoom=15, long=NA, lat=NA) 
  
  reactive({
    # this will redraw the leaflet
    cat(file=stderr(), paste('initialise map centre'), "\n")
    v$long <- isolate(my$long)
    v$lat <- isolate(my$lat)
  })
  
  # observeEvent(input$zoomin, {
  #   v$zoom <- ifelse(v$zoom<v$maxzoom, v$zoom+1, v$zoom)
  # }) # end observeEvent
  # 
  # observeEvent(input$zoomout, {
  #   v$zoom <- ifelse(v$zoom>v$minzoom, v$zoom-1, v$zoom)
  # }) # end observeEvent
  
  # titlePanel("Where are You?")
  output$seasonSelector <- renderUI({
    cat(file=stderr(), paste('render season selector'), "\n")
    selectInput("season", h4("Which Season?"), as.list(seasons), selected=my$season)
  })
  
  # Make your initial map
  # https://stackoverflow.com/questions/34348737/r-leaflet-how-to-click-on-map-and-add-a-circle
  output$map <- renderLeaflet({
    cat(file=stderr(), paste('render leaflet'), "\n")
    
      isolate({ # prevent redraw if arguments change
        leaflet(spgons, options=leafletOptions(minZoom=v$minzoom, maxZoom=v$maxzoom)) %>%
          setView(v$long, v$lat, zoom=v$zoom)  %>%
          addTiles() %>% # default map
          addPolygons(data=spgons, color=spgon_cols, weight=0, options=pathOptions(clickable=FALSE)) %>%
          addMarkers(my$long, my$lat, 'layer1') %>%
          addCircles(my$long, my$lat, layerId=as.character(windows), radius=windows*1000, 
                     color=window_cols, weight=4, fill=NA) 
          # addCircles(my$long, my$lat, 'layer2', radius=windows[1]*1000, weight=2, fill=NA) %>%
          # addCircles(my$long, my$lat, 'layer3', radius=windows[2]*1000, weight=2, fill=NA) %>%
          # addCircles(my$long, my$lat, 'layer4', radius=windows[3]*1000, weight=2, fill=NA)
      })

    }) # end renderLeaflet
  
  #### define ui ####
  sidebarLayout(
    
    sidebarPanel(
      uiOutput("seasonSelector") # this control is created in the server
      # actionButton("go", "Go!")
    ), # end sidebarPanel
    
    mainPanel(
      leafletOutput("map")
    ) # end mainPanel
    
  ) # end sidebarLayout

  # Observe mouse clicks 
  # see also https://rstudio.github.io/leaflet/shiny.html
  observeEvent(input$map_click, {
    cat(file=stderr(), paste('map_click'), "\n")    

    click <- input$map_click
    clat <- click$lat
    clng <- click$lng
    address <- revgeocode(c(clng,clat))
    my$long <- clng
    my$lat <- clat
    # v$long <- input$map_lng
    # v$lat <- input$map_lat
    # v$zoom <- input$map_zoom
    # 
    leafletProxy('map') %>%
      addMarkers(my$long, my$lat, 'layer1') %>%
      addCircles(my$long, my$lat, layerId=as.character(windows), radius=windows*1000, color=window_cols, weight=4, fill=NA) 
    
  })

  # this ignores NULL values 
  observeEvent(input$season, { 
    cat(file=stderr(), paste('season selection'), "\n")
    my$season <- input$season
  })
  
```

<!-- # Your Neighbourhood -->

```{r eval=TRUE, echo=FALSE, warning=TRUE}

calc <- reactive({

    req(input$season) # this prevents it running before input$season is defined
    
    # reactive dependence on these variables (they triggering recalc)
    # my$long # triggered by "map" _click
    # my$lat # triggered by "map" _click
    #my$season <- input$season #  FIXME causing problems because input$season not initialised!

    my$name <- paste('You (', as.character(my$season), ')', sep='')
  
    cat(file=stderr(), paste('my$long my$lat =', my$long, my$lat), "\n")
    cat(file=stderr(), paste('my$season =', my$season), "\n")

  # filer by season. 
  # Warning: What if there is insufficient data!!!!
  data_sel <-  data_all %>% filter(season==my$season) 

  # location for map centre
  nzgd <- data.matrix(c(my$long, my$lat))
  nztm <- project(xy=nzgd, proj=proj4string)
  my$east <- nztm[,1]
  my$north <- nztm[,2]

  # calculate aspect ratio near my farm
  nzgd <- data.matrix(tibble(long=c(my$long, my$long, my$long-0.5, my$long+0.5),
                             lat=c(my$lat-0.5, my$lat+0.5, my$lat, my$lat)))
  nztm <- project(xy=nzgd, proj=proj4string)
  my$distortion <- (max(nztm[,2])-min(nztm[,2]))/(max(nztm[,1])-min(nztm[,1]))

  # calculate distances ####
  # we need to use rowwise()  because distm is not vectorised, I think, although rowwise() is deprecated
  # http://www.expressivecode.org/2014/12/17/mutating-using-functions-in-dplyr/
  data_sel <- data_sel %>% 
    rowwise() %>% 
    mutate(
         dist = distm(c(my$long, my$lat), c(long, lat), fun=distHaversine), # this needs rowwise()
         dist = dist/1000 # km
  )

  # filter by distance
  data_sel <- data_sel %>% filter(dist < max(windows))

  # empty data frames for loop
  farms <- tibble(x=numeric(), y=numeric(), east=numeric(), north=numeric(), long=numeric(), lat=numeric(),
                  pasture=numeric(), dist=numeric(), window=numeric(), radius=factor())
  sampcdf <- tibble(probs=numeric(), quants=numeric(), radius=factor())
  samppdf <- tibble(pasture=numeric(), window=numeric(), radius=factor(),
                    q=numeric(), qr=numeric(), qrlower=numeric(), qrupper=numeric())
  circles <- tibble(east=numeric(), north=numeric(), radius=factor())
  
  # circle function
  circle_fun <- function(centre=c(0,0), r=1, npoints=100){
    tt <- seq(0, 2*pi, length.out=npoints)
    xx <- centre[1] + r * cos(tt)
    yy <- centre[2] + r * sin(tt)
    return(tibble(x=xx, y=yy))
  }    
    
  # loop through decreasing window sizes 
  for (window in windows) {

    # select data within window
    data_window <- data_sel %>% filter(dist < window)
    n <- nrow(data_window)
    code <- paste(window,' km (',n,' Farms)', sep='')

    # calculate circle
    circle <- circle_fun(centre=c(my$east, my$north), r=window*1000, npoints=100)
    nztm <- data.matrix(circle[,c('x', 'y')])
    nzgd <- project(xy=nztm, proj=proj4string, inverse=TRUE)
    circle$long <- nzgd[,1]
    circle$lat <- nzgd[,2]

    # save selected farms for plot
    if (n >= 1) {
      farms <- rbind(farms, tibble(east=data_window$east, north=data_window$north,
                                 long=data_window$long, lat=data_window$lat,
                                 pasture=data_window$pasture_eaten,
                                 dist=data_window$dist, window=window, radius=as.factor(code)))
    }
    
    circles <- rbind(circles, tibble(long=circle$long, lat=circle$lat, radius=as.factor(code)))

    # calculate quantile
#    qr1 <- rq(formula=pasture_eaten ~ east + north, tau=0.9, data=data_window) # linear quantile regression
    qr1 <- rq(formula=pasture_eaten ~ 1, tau=0.9, data=data_window) # linear quantile regression
    se_method <- "boot" # how condience intervals are calculated, some methods more robust
    yqr1<- predict(qr1, tibble(east=my$east, north=my$north), interval='confidence', level=0.95, se=se_method)
    q90 <- quantile(data_window$pasture_eaten, 0.9, type=1) # also calc simple q90

    cat(file=stderr(), paste('yqr1 =', yqr1), "\n")
    cat(file=stderr(), paste('q90 =', q90), "\n")

    # save sample quantiles if enough data to be sensible
    if (n >= minn) {
      quants <- quantile(data_window$pasture_eaten, probs=probs, type=8) # see documentation for type=?
      sampcdf <- rbind(sampcdf, tibble(probs=probs, quants=quants, radius=as.factor(code)))
      samppdf <- rbind(samppdf, tibble(pasture=data_window$pasture_eaten, window=window, radius=as.factor(code),
                       q=q90, qr=yqr1[1], qrlower=yqr1[2], qrupper=yqr1[3]))
    }

  } # next window size
    
#   # little function to return a quantile
#   qfn <- function(x){
# #    q <- quantile(x, 0.9, na.rm=TRUE)
#     q <- quantile(x, 0.9)
#     return(q)
#   }
# 
#   # add quantiles to data
#   samppdf <- samppdf %>%
#     group_by(radius) %>%
#     mutate(q=qfn(pasture))

  # biggest circle
  circle <- circle_fun(centre=c(my$east, my$north), r=max(windows)*1000, npoints=100)
  nztm <- data.matrix(circle[,c('x', 'y')])
  nzgd <- project(xy=nztm, proj=proj4string, inverse=TRUE)
  circle$long <- nzgd[,1]
  circle$lat <- nzgd[,2]

  # return results
  return(list(data_sel=data_sel, circles=circles, circle=circle, farms=farms, sampcdf=sampcdf, samppdf=samppdf))
  
})

```



```{r eval=FALSE, echo=FALSE}
# for testing
  renderTable({ head(calc()$data_sel) })
  renderTable({ reactiveValuesToList(my) }) # my is not a data table
  renderTable({ head(calc()$circles) })
  renderTable({ head(calc()$farms) })
  renderTable({ head(calc()$sampcdf) })
  renderTable({ head(calc()$samppdf) })

```

<!-- This is the area you have selected. The nearby farms which have data are shown on the second map. -->

```{r eval=FALSE, echo=FALSE}

# nz map
# output$plot1 <- renderPlot({
# 
#   data_sel <- calc()$data_sel
#   circle <- calc()$circle
#   title <- paste('Farm Locations near', my$name)
#   title <- 'Location'
#   
#   # plot nearby locations on a map
#   plot1 <- ggplot() +
#     labs(title=title, x='Longitude', y='Latitude', colour="Region") +
# #    geom_point(data=data_sel, mapping=aes(x=long,y=lat,colour=region),shape=16,size=2) +
#     scale_colour_manual(values=cb9sron) +
#     coord_fixed() + # fix aspect ratio
#     theme_cowplot() +
#     geom_path(data=nz, mapping=aes(x=long,y=lat,group=group), colour='grey') +
#     geom_path(data=circle, mapping=aes(x=long,y=lat), colour='red') +
#     coord_map() +
#     geom_point(mapping=aes(x=my$long,y=my$lat),shape=4,size=2,stroke=2,colour='black') +
#     panel_border(colour='black')
# 
#     plot1
# 
# })

# local googlemap
output$plot2 <- renderPlot({

  circles <- calc()$circles
  farms <- calc()$farms
  my_temp <- tibble(long=my$long, lat=my$lat) # avoid error casued by using my in geom_point()
  
  # get googlemap
  # google map
  # my_map <- get_googlemap(center=c(175.352116, lat=-37.781841), zoom=9, scale=2, maptype='hybrid') 
  my_map <- get_googlemap(center=c(my$long, my$lat), zoom=9, scale=2, maptype='hybrid') 
  plot2 <- ggmap(my_map, extent='panel', darken=c(0.4, 'white'))  # 'normal' adds a white border
  
  title <- paste('Farm Locations near', my$name)
  title <- 'Map'
  
  # construct area plot
  plot2 <- plot2 +
    labs(title=title, x='East', y='North',
         colour="Radius (km)", size="Pasture Eaten") +
    coord_fixed(ratio=my$distortion) + # fix aspect ratio
    theme_cowplot() +
    panel_border(colour='black')  

    plot2 <- plot2 +
      geom_polygon(data=spgonsdf, aes(y=lat, x=long, group=group, fill=id)) +
      scale_fill_manual(values=spgon_cols) +
      guides(fill=FALSE)

    # this works
    # ggplot(data=spgons_df, aes(y=lat, x=long, group=group, fill=id)) +
    # geom_polygon() +
    # guides(fill=FALSE)
  
  #  area plot
  plot2 <- plot2 +
    geom_path(data=circles, mapping=aes(x=long, y=lat, group=radius, colour=radius), size=1.2) +
#    geom_point(data=farms, mapping=aes(x=long, y=lat, size=pasture, group=radius, colour=radius)) +
    geom_point(data=my_temp, mapping=aes(x=long, y=lat), shape=4, size=2, stroke=2, colour='black') +
    theme(legend.position="bottom", legend.box='vertical') 

  plot2

})


fluidPage({
  fluidRow(
      # column(width=4, 
      #   plotOutput("plot1")
      # ), # end column
      column(width=8, 
        plotOutput("plot2")
      ) # end column
    ) # end fluidRow
})

```

This is the distribution of pasture produced and consumed within various distances of your location. The 90th percentile is also shown.

```{r eval=TRUE, echo=FALSE}

# stacked histograms
output$plot7 <- renderPlot({

  samppdf <- calc()$samppdf

  breaks <- seq(floor(min(samppdf$pasture))-1, ceiling(max(samppdf$pasture))+1, 1)

  cat(file=stderr(), paste('xlim =', min(breaks), max(breaks)), "\n")
  
  title_string <- paste('Pasture Eaten near', my$name)
  
  plot7 <- ggplot() +
    labs(title=title_string, y='Number of Farms', x='Pasture Eaten (tDM '*ha^-1~y^-1*')', colour='Radius (km)') +
    theme_cowplot() +
  #  scale_y_continuous(breaks=c()) + # remove y-scale when too many facets
    panel_border(colour='black') +  
    theme(legend.position='none')
  
  if (nrow(samppdf)>0) { 
  plot7 <- plot7 +
    geom_rect(data=samppdf, mapping=aes(xmin=qrlower, xmax=qrupper, ymin=0, ymax=Inf), 
              fill='lightcyan') +
    geom_histogram(data=samppdf, mapping=aes(x=pasture, colour=radius), 
                   fill=NA, size=1.1, binwidth=1) +
    geom_vline(data=samppdf, mapping=aes(xintercept=qr), size=1.5, colour='lightcyan4', alpha=0.2) +
    geom_vline(data=samppdf, mapping=aes(xintercept=q), size=1.5, colour='black') +
    geom_text(data=samppdf, mapping=aes(x=q, y=4, label='90th'), hjust=0, nudge_x=0.1) +
    facet_grid(radius ~ ., as.table=FALSE) + # as.table=FALSE reverses the order
    theme(strip.background=element_blank(), strip.text.y=element_text(angle=0)) +
    scale_x_continuous(breaks=breaks) +
    coord_cartesian(xlim=c(min(breaks),max(breaks)))

  }
  
  plot7

}) # end renderPlot
  
fluidPage({
    fluidRow(
      plotOutput("plot7")
    ) # end fluidRow
})

```

