---
title: "Potential Pasture and Crop Eaten"
runtime: shiny
output: 
  prettydoc::html_pretty:
    theme: architect
---

<!-- Simon Woodward, DairyNZ, 2019 -->
<!-- now use joined.rds data, including elevation nitrogen etc -->
<!-- # Note: Unlike Shiny Apps, Interactive R Markdown Documents are do not require ui and server,  -->
<!-- #       the whole document is treated as a server. -->
<!-- # You can't guarantee what order the reactive elements will execute -->


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)

# https://shiny.rstudio.com/articles/shinyapps.html
# note: shiny attempts to upload only packages which are actually used

library(tidyverse)
library(leaflet) # interactive map
library(ggmap) # register google api
library(cowplot) # plot box
library(geosphere) # distm
library(proj4) # warning masked by rgdal
library(KernSmooth) # for contouring
library(sp) # spatial polygons
library(quantreg) # quantile regression
library(ggthemes) # chart theme
library(dismo) # geocode
library(raster) # required by geocode. Warning - masks dplyr::select
library(XML) # needed by geocode
library(googleway) # for geocoding

# https://stackoverflow.com/questions/36175529/getting-over-query-limit-after-one-request-with-geocode
# https://developers.google.com/maps/documentation/geocoding/get-api-key
# https://lucidmanager.org/geocoding-with-ggmap/
googlekey <- readLines("google.api") # text file with the API key
# ggmap::register_google(key = googlekey)
# ggmap::ggmap_credentials()

#### these objects are available to all sessions ####
# this is good place to put global data or utility functions
# you can change them (for every session) but you have to use the <<- operator

# read multiyear data and tidy up 
if (FALSE){ # jitter data
  data_all <- readRDS("joined.rds") %>%
    dplyr::select(pasture_and_crop_eaten_t_dm_ha, region, season, long, lat,
                  nzsc_order, elev,
                  nitrogen_applied_for_year_l2) %>% 
    tidyr::drop_na() # drops a lot of rows!
  data_all <- data_all %>% # jitter
    plyr::mutate(long=long+0.1*(runif(nrow(data_all))-runif(nrow(data_all))),
                 lat=lat+0.1*(runif(nrow(data_all))-runif(nrow(data_all)))
                 )
  saveRDS(data_all, "joined_comp.rds")
}
data_all <- readRDS("joined_comp.rds") %>%
  plyr::mutate(pasture_eaten=as.numeric(pasture_and_crop_eaten_t_dm_ha),
         region=as.factor(region),
         season=as.factor(season),
         nitrogen_applied=as.numeric(nitrogen_applied_for_year_l2),
         soil=as.character(nzsc_order))  
temp <- nrow(data_all)
data_all <- data_all %>% tidyr::drop_na() # drops a lot of rows!
cat(file=stderr(), paste("Rows dropped with missing data =", temp-nrow(data_all), "\n"))
cat(file=stderr(), paste("Rows remaining =", nrow(data_all), "\n"))

# add elevation factor
data_all <- data_all %>%
  mutate(elev_fact=factor(x=case_when(elev<=50 ~ "Low (0-50m)", 
                                      elev<=200 ~ "Middle (50-200m)", 
                                     TRUE ~ "High (200m+)"),
                          levels=c("Low (0-50m)", "Middle (50-200m)", "High (200m+)")))

# summary(data_all)

# create season and soil list
season_all <- as.list(sort(unique(as.character(data_all$season))))
names(season_all) <- season_all
soil_all <- as.list(sort(unique(data_all$soil)))
names(soil_all) <- soil_all
names(soil_all)[soil_all=="L"] <- "Allophanic"
names(soil_all)[soil_all=="A"] <- "Anthropic"
names(soil_all)[soil_all=="B"] <- "Brown"
names(soil_all)[soil_all=="Z"] <- "Podzol"
names(soil_all)[soil_all=="M"] <- "Pumice"
names(soil_all)[soil_all=="W"] <- "Raw"
names(soil_all)[soil_all=="G"] <- "Gley"
names(soil_all)[soil_all=="N"] <- "Granular"
names(soil_all)[soil_all=="E"] <- "Melanic"
names(soil_all)[soil_all=="R"] <- "Recent"
names(soil_all)[soil_all=="S"] <- "Semiarid"
names(soil_all)[soil_all=="U"] <- "Ultic"
names(soil_all)[soil_all=="O"] <- "Organic"
names(soil_all)[soil_all=="X"] <- "Oxidic"
names(soil_all)[soil_all=="P"] <- "Pallic"
elev_all <- levels(data_all$elev_fact)
names(elev_all) <- elev_all

# define nitrogen steps
nitrogen_range <- seq(0L, 250L, 50L)
nitrogen_here <- as.list(c("Don't Adjust", paste(nitrogen_range, "kgN/ha/y")))
nitrogen_default <- nitrogen_here[[4]]

# calculate nitrogen adjustment
data_all <- data_all %>%
  mutate(
    nitrogen_slope = 0.010,
    pasture_eaten_raw = pasture_eaten,
    pasture_eaten_min = pasture_eaten + (min(nitrogen_range) - nitrogen_applied) * nitrogen_slope,
    pasture_eaten_max = pasture_eaten + (max(nitrogen_range) - nitrogen_applied) * nitrogen_slope
    )

# create data point map/contours for plotting on leaflet
# https://gis.stackexchange.com/questions/168886/r-how-to-build-heatmap-with-the-leaflet-package
data_pts <- unique(data_all[c("long", "lat")])
kde <- KernSmooth::bkde2D(data.matrix(data_pts), bandwidth=c(0.1, 0.1), gridsize=c(100,100))
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat) # contour lines (list)
LEVS <- as.factor(sapply(CL, `[[`, "level")) # contour levels (vector)
NLEV <- length(levels(LEVS)) # number of levels 
pgons <- lapply(1:length(CL), function(i) # convert to polygons (ID=i is actually the level)
    sp::Polygons(list(sp::Polygon(cbind(CL[[i]]$x, CL[[i]]$y))), ID=i))
spgons = sp::SpatialPolygons(pgons)
spgon_cols <- topo.colors(NLEV, NULL)[LEVS]

# calculate NZTM2000 coordinates for farm locations
proj4string <- "+proj=tmerc +lat_0=0.0 +lon_0=173.0 +k=0.9996 +x_0=1600000.0 +y_0=10000000.0 +datum=WGS84 +units=m"
nzgd <- data.matrix(data_all[,c("long", "lat")])
nztm <- proj4::project(xy=nzgd, proj=proj4string)
temp <- proj4::project(xy=nzgd, proj=proj4string, inverse=TRUE)
data_all$east <- nztm[,1]
data_all$north <- nztm[,2]

# define some constants 
trim <- 0.0 # rqss fails near tails if insufficient data
probs <- seq(trim, 1-trim, 0.02) # probabilities for sampcdf
nprobs <- length(probs)
windows <- c(60,40,20)
nmin <- 4L # minimum number of farms in a window for analysis

# gets a list of default ggplot colours
gg_colour_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
window_cols <- gg_colour_hue(length(windows))

# default_loc must be somewhere with data!!!
# default_loc <- list(long=172.833333, lat=-41.5) # Nelson
default_loc <- list(long=175.619105, lat=-40.386396) # Massey
# default_loc <- list(long=174.865530, lat=-41.259256) # Wellington
# default_loc <- list(long=175.352116, lat=-37.781841) # DairyNZ
# default_loc <- list(long=167.893066, lat=-43.036775) # In the ocean!

min_long <- 160
max_long <- 180
min_lat <- -50
max_lat <- -30
# default_location <- "Click map or enter location here"
default_location <- ""

# normal list with same elements, useful for testing
my <- list(name="You Are Here",
                     long=default_loc$long,
                     lat=default_loc$lat,
                     distortion=1,
                     east=NA, north=NA,
                     adjust_here=nitrogen_here,
                     adjust_default=nitrogen_default,
                     adjust=nitrogen_default,
                     season_here=list(NA),
                     season=list(NA),
                     soil_here=list(NA),
                     soil=list(NA),
                     elev_here=list(NA),
                     elev=list(NA),
                     data=list(NA),
                     breaks=list(NA),
                     recalc=0L
                     )

```

Pasture and crop eaten is the fundamental basis of profitable dairy systems. In general, the more pasture and crop eaten, the more profitable a farm will be – an extra tonne of DM eaten per hectare is associated with [over $300](https://www.dairynz.co.nz/media/5747949/pasture-eaten.png) per hectare of extra operating profit. This tool allows you to determine what the pasture and crop eaten is on other farms in your region, and to consider what the potential might be. You can then compare this to the amount eaten on your farm (e.g. from a [DairyBase report](https://www.dairynz.co.nz/business/dairybase/)) or by calculating it yourself with [these tools](https://www.dairynz.co.nz/feed/pasture-management/assessing-farm-performance/pasture-and-crop-eaten). This can indicate whether there is a gap between potential and what you are currently achieving, and hence if there may be an opportunity to increase pasture and crop eaten and profitability.

#### Where are You?

Please click your location on the map. The coloured contours on the map show where data availability is higher, and so information will be more meaningful. Then select which season, [soil order(s)](https://soils.landcareresearch.co.nz/describing-soils/nzsc/soil-order/) and altitude(s) of data to include. You can also adjust the pasture and crop eaten data to reflect difference in nitrogen fertiliser use to match your nitrogen application rate. 

```{css, echo = FALSE}
/* fix cursor */
.leaflet-container {
  cursor: auto !important;
}
```

```{r input_pane, echo=FALSE, eval=TRUE}

#### initialise session ####

# collect info about the current location and selections
# these variables are available in the reactive conext
# "my" is not itself a reactive object, it's a list of reactive objects
my <- reactiveValues(name="You Are Here", 
                     long=default_loc$long,
                     lat=default_loc$lat,
                     distortion=1, 
                     east=NA, north=NA, 
                     adjust_here=nitrogen_here,
                     adjust_default=nitrogen_default,
                     adjust=nitrogen_default,
                     season_here=list(NA), 
                     season_default=list(NA), 
                     season=list(NA),
                     soil_here=list(NA), 
                     soil_default=list(NA), 
                     soil=list(NA),
                     elev_here=list(NA), 
                     elev_default=list(NA), 
                     elev=list(NA),
                     data=list(NA),
                     breaks=list(NA),
                     recalc=0L
                     )

# set leaflet info
v <- reactiveValues(zoom=5, minzoom=5, maxzoom=15, long=NA, lat=NA) 

# initialise map centre
isolate({
  cat(file=stderr(), paste("initialise location"), "\n")
  v$long <- my$long
  v$lat <- my$lat
})

#### ui elements ### 

# titlePanel("Where are You?")

output$season_selector <- renderUI({
  cat(file=stderr(), paste("render season selector"), "\n")
  selectInput("season", strong("Production season?"), my$season_here, selected=my$season_default)
})

output$elev_selector <- renderUI({
  cat(file=stderr(), paste("render elev selector"), "\n")
  selectInput("elev", strong("Altitude over sea level?"), my$elev_here, selected=my$elev_default,
              selectize=FALSE, multiple=TRUE)
})

output$soil_selector <- renderUI({
  cat(file=stderr(), paste("render soil selector"), "\n")
  selectInput("soil", strong("Soil orders?"), my$soil_here, selected=my$soil_default, 
              selectize=FALSE, multiple=TRUE)
})

output$nitrogen_checkbox <- renderUI({
  cat(file=stderr(), paste("render adjust checkbox"), "\n")
  selectInput("adjust", strong("Nitrogen fertiliser applied?"), my$adjust_here, selected=my$adjust_default)
})

#### make initial map ####

# https://stackoverflow.com/questions/34348737/r-leaflet-how-to-click-on-map-and-add-a-circle
output$map <- renderLeaflet({
  cat(file=stderr(), paste("render leaflet"), "\n")
  
    isolate({ # prevent redraw if arguments change

      leaflet(spgons, options=leafletOptions(minZoom=v$minzoom, maxZoom=v$maxzoom)) %>%
        setView(v$long, v$lat, zoom=v$zoom)  %>%
        addTiles() %>% # default map
        addPolygons(data=spgons, color=spgon_cols, weight=0, options=pathOptions(clickable=FALSE)) %>%
        addMarkers(my$long, my$lat, "layer1", options=pathOptions(clickable=FALSE)) %>%
        addCircles(my$long, my$lat, layerId=as.character(windows), radius=windows*1000, 
                   color=window_cols, weight=4, fill=NA, options=pathOptions(clickable=FALSE))
      
    })

  }) # end renderLeaflet

#### ui layout ####

css <- "
.selectize-input { 
font-size: 14px; line-height: 14px;
} 
.selectize-dropdown { 
font-size: 14px; line-height: 14px; 
}"

shinyUI(fluidPage(

  tags$style(type='text/css', css),

  sidebarLayout(
    
    sidebarPanel(
      cat(file=stderr(), paste("render sidebar"), "\n"),
      width=5, # 12ths of the panel
      uiOutput("season_selector"), 
      uiOutput("elev_selector"),
      uiOutput("soil_selector"), 
      uiOutput("nitrogen_checkbox") 
    ), # end sidebarPanel
    
    mainPanel(
      cat(file=stderr(), paste("render main panel"), "\n"),
      width=7, # 12ths of the panel
      leafletOutput("map", width="100%", height=480), # can manipulate size here
      absolutePanel(top=10, left=70, textInput("search_bar", "" , default_location, "75%")) # search bar
    ), # end mainPanel
    
    position="right"
    
  ) # end sidebarLayout

)) # end fluidPage

#### react to mouse clicks ####

# see also https://rstudio.github.io/leaflet/shiny.html
observeEvent(input$map_click, {
  
  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed map_click"), "\n")    

  click <- input$map_click
  my$long <- click$lng
  my$lat <- click$lat
  my$season <- list(NA)
  my$elev <- list(NA)
  my$soil <- list(NA)
  
  # update map
  leafletProxy("map", deferUntilFlush=FALSE) %>%
    addMarkers(my$long, my$lat, "layer1", options=pathOptions(clickable=FALSE)) %>%
    addCircles(my$long, my$lat, layerId=as.character(windows), radius=windows*1000, 
               color=window_cols, weight=4, fill=NA, options=pathOptions(clickable=FALSE)) 
  
}) # end observe mouse click

# https://www.r-bloggers.com/4-tricks-for-working-with-r-leaflet-and-shiny/
observeEvent(input$search_bar, {

  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed search bar =", input$search_bar), "\n")
  req(input$search_bar, nchar(input$search_bar)>=4)
  req(googlekey>"")
  
  # target_pos <- dismo::geocode(paste(input$search_bar, "New Zealand"),
  #                              extent=raster::extent(min_long, max_long, min_lat, max_lat),
  #                              oneRecord=TRUE)
  target_pos <- googleway::google_geocode(paste(input$search_bar, "New Zealand"),
                                          bounds=list(c(min_lat,min_long),c(max_lat,max_long)),
                                          key=googlekey)$results$geometry$location %>% 
    rename(lon=lng)
  cat(file=stderr(), paste(input$search_bar, "=", target_pos$lon, target_pos$lat), "\n")
  req(target_pos$lon, target_pos$lat)
  req(target_pos$lon!="NA", target_pos$lat!="NA")
  req(target_pos$lon>=min_long, target_pos$lon<=max_long)
  req(target_pos$lat>=min_lat, target_pos$lat<=max_lat)
    
  my$long <- target_pos$lon
  my$lat <- target_pos$lat
  my$season <- list(NA)
  my$elev <- list(NA)
  my$soil <- list(NA)

  # update map
  leafletProxy("map", deferUntilFlush=FALSE) %>%
    setView(my$long, my$lat, zoom=input$map_zoom)  %>%
    addMarkers(my$long, my$lat, "layer1", options=pathOptions(clickable=FALSE)) %>%
    addCircles(my$long, my$lat, layerId=as.character(windows), radius=windows*1000,
               color=window_cols, weight=4, fill=NA, options=pathOptions(clickable=FALSE))

}) # end observe search bar
          
#### react to change of location ####

observeEvent(c(my$long, my$lat), {

  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed location =", my$long, my$lat), "\n")
    
  # calculate aspect ratio near my farm (not used)
  # nzgd <- data.matrix(tibble(long=c(my$long, my$long, my$long-0.5, my$long+0.5),
  #                            lat=c(my$lat-0.5, my$lat+0.5, my$lat, my$lat)))
  # nztm <- proj4::project(xy=nzgd, proj=proj4string)
  # my$distortion <- (max(nztm[,2])-min(nztm[,2]))/(max(nztm[,1])-min(nztm[,1]))
  
  # location for map centre
  nzgd <- data.matrix(c(my$long, my$lat))
  nztm <- proj4::project(xy=nzgd, proj=proj4string)
  my$east <- nztm[,1]
  my$north <- nztm[,2]

  # calculate distance #
  # we need to use rowwise()  because distm is not vectorised, although rowwise() is deprecated
  # ungroup() removes the effect of rowwise()
  # http://www.expressivecode.org/2014/12/17/mutating-using-functions-in-dplyr/
  data <- data_all %>% 
    rowwise() %>% 
    mutate(dist = geosphere::distm(c(my$long, my$lat), c(long, lat), fun=distHaversine), # this needs rowwise()
           dist = dist/1000) %>%
    ungroup()
  
  # filter data
  temp <- data %>% 
    filter(dist < max(windows)) # km
  
  # insufficient data?
  if (nrow(temp)>=nmin){
    data <- temp
  } else {
    data <- data[1:max(2,nmin-1),] # grab a few rows (they won't be shown)
  }

  # calculate width of histogram for region
  my$breaks <- seq(  floor(min(data$pasture_eaten_raw, data$pasture_eaten_min, data$pasture_eaten_max)), 
                   ceiling(max(data$pasture_eaten_raw, data$pasture_eaten_min, data$pasture_eaten_max)), 
                   1)
  
  # store data for region
  my$data <- data 
 
  # what season are available here   
  i <- sort(unique(as.character(data$season)))
  my$season_here <- season_all[match(i, season_all)]
  n <- unlist(map(my$season_here, function(u) sum(u==data$season)))
  names(my$season_here) <- paste(names(my$season_here), " (", n, " Farms)", sep="")
  cat(file=stderr(), paste("my$season_here =", length(my$season_here)), "\n")
  cat(file=stderr(), paste(names(my$season_here)), "\n")

  # reset selections
  my$season_default <- tail(my$season_here, 1)
  my$season <- my$season_default
  my$elev <- list(NA)
  my$soil <- list(NA)
  # don't change my$adjust
  
}) # end reaction to location changing

#### react to change of season ####

observeEvent(input$season, {
  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed input$season =", input$season), "\n")
  req(input$season, input$season!="NA")
  if (my$season != input$season) {
    my$season <- input$season
    my$elev <- list(NA)
    my$soil <- list(NA)
    # don't change my$adjust
  }
})

observeEvent(my$season, {

  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed my$season =", my$season), "\n")
  req(my$season, my$season!="NA")

  # make elev list
  data <- my$data %>%
    filter(season == my$season)
  
  # insufficient data?
  if (nrow(data)<nmin){
    data <- my$data[1:max(2,nmin-1),]
  }
  
  # data <- data_all[1:10,] # subset for testing
  i <- unique(data$elev_fact)
  my$elev_here <- elev_all[which(elev_all %in% i)] # retains sorting
  n <- unlist(map(my$elev_here, function(u) sum(u==data$elev_fact)))
  names(my$elev_here) <- paste(my$elev_here, " (", n, " Farms)", sep="")
  cat(file=stderr(), paste("my$elev_here =", length(my$elev_here)), "\n")
  cat(file=stderr(), paste(names(my$elev_here)), "\n")

  # reset selections
  my$elev_default <- my$elev_here
  my$elev <- my$elev_default 
  my$soil <- list(NA)
  # don't change my$adjust
  
}) # end reaction to season changing

#### react to change of elevation ####

observeEvent(input$elev,  {
  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed input$elev ="), paste(input$elev), "\n")
  req(input$elev, input$elev!="NA")
  if (any(my$elev != input$elev, na.rm=TRUE)) {
    my$elev <-  input$elev
    my$soil <- list(NA)
    # don't change my$adjust
  }
})

observeEvent(my$elev,  {

  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed my$elev ="), paste(my$elev), "\n")
  req(my$elev, my$elev!="NA")

  # make soil list
  data <- my$data %>%
    filter(season == my$season) %>%
    filter(elev_fact %in% my$elev) 
  
  # insufficient data?
  if (nrow(data)<nmin){
    data <- my$data[1:max(2,nmin-1),]
  }

  i <- sort(unique(data$soil))
  my$soil_here <- soil_all[match(i, soil_all)]
  n <- unlist(map(my$soil_here, function(u) sum(u==data$soil)))
  names(my$soil_here) <- paste(names(my$soil_here), " (", n, " Farms)", sep="")
  cat(file=stderr(), paste("my$soil_here =", length(my$soil_here)), "\n")
  cat(file=stderr(), paste(names(my$soil_here)), "\n")

  # reset selections
  my$soil_default <- my$soil_here
  my$soil <- my$soil_default 
  # don't change my$adjust
  
})

#### react to change of soil ####

observeEvent(input$soil, {
  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed input$soil ="), paste(input$soil), "\n")
  req(input$soil, input$soil!="NA")
  if (any(my$soil != input$soil, na.rm=TRUE)) {
    my$soil <- input$soil
    # don't change my$adjust
  }
})

observeEvent(my$soil, {
  
  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed my$soil = "), paste(my$soil), "\n")
  req(my$soil, my$soil!="NA")
  
  # trigger recalc
  my$recalc <- my$recalc + 1L
  
}) # end reaction to soil changing

#### react to change of adjust ####

observeEvent(input$adjust, {
  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed input$adjust =", input$adjust), "\n")
  req(my$adjust, input$adjust)
  if (my$adjust != input$adjust) {
    my$adjust <- input$adjust
  }
})

observeEvent(my$adjust, {
  
  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("observed my$adjust =", my$adjust), "\n")
  req(my$adjust)

  # trigger recalc
  my$recalc <- my$recalc + 1L
  
})

```

<!-- # Your Neighbourhood -->

```{r output_calculations, eval=TRUE, echo=FALSE, warning=TRUE}

#### react to change of inputs ####

calc <- eventReactive(my$recalc,  {

  cat(file=stderr(), "\n")
  cat(file=stderr(), paste("analyse for location =", my$long, my$lat), "\n")
  cat(file=stderr(), paste("analyse for season =", my$season), "\n")
  cat(file=stderr(), paste("analyse for soil ="), paste(my$soil), "\n")
  cat(file=stderr(), paste("analyse for elev ="), paste(my$elev), "\n")
  cat(file=stderr(), paste("analyse for adjust ="), my$adjust, "\n")
  req(my$season!="NA", my$soil!="NA", my$elev!="NA", my$adjust!="NA")
  req(my$long, my$lat, my$season, my$soil, my$elev, my$adjust)

  dont_adjust <- nitrogen_here[[1]]
  nitrogen_level <- as.numeric(word(my$adjust,1))
  data <-  my$data %>% 
    filter(season == my$season) %>% 
    filter(soil %in% my$soil) %>%
    filter(elev_fact %in% my$elev) %>%
    mutate(pasture_eaten = 
             case_when(
               my$adjust==dont_adjust ~ pasture_eaten_raw,
               TRUE ~ pasture_eaten_raw + (nitrogen_level - nitrogen_applied) * nitrogen_slope
               )
           )

  if (my$adjust==dont_adjust){
    my$name <- paste("Your Location (", season_all[my$season[[1]]], ")", sep="")
  } else {
    my$name <- paste("Your Location (", season_all[my$season[[1]]], ")",
                     # " (Nitrogen response = ", sprintf("%.1f", data$nitrogen_slope[1]*1000), ")", 
                     sep="")
  }
  
  cat(file=stderr(), paste("nrow(data) = ", nrow(data)), "\n")

  # circle function
  circle_fun <- function(centre=c(0,0), r=1, npoints=100){
    tt <- seq(0, 2*pi, length.out=npoints)
    xx <- centre[1] + r * cos(tt)
    yy <- centre[2] + r * sin(tt)
    return(tibble(x=xx, y=yy))
  }    
    
  # prepare empty data frames for loop
  farms <- tibble(x=numeric(), y=numeric(), east=numeric(), north=numeric(), long=numeric(), lat=numeric(),
                  pasture=numeric(), dist=numeric(), window=numeric(), radius=factor())
  sampcdf <- tibble(probs=numeric(), quants=numeric(), radius=factor())
  samppdf <- tibble(pasture=numeric(), window=numeric(), radius=factor(),
                    q=numeric(), qr=numeric(), qrlower=numeric(), qrupper=numeric())
  circles <- tibble(east=numeric(), north=numeric(), radius=factor())
  
  # loop through decreasing window sizes 
  for (window in windows) {

    # select data within window
    data_window <- data %>% filter(dist < window)
    n <- nrow(data_window)
    code <- paste(window," km\n(", format(n, width=1), " Farms)", sep="")
    cat(file=stderr(), paste("window = ", window, " km"), "\n")

    # calculate circle
    circle <- circle_fun(centre=c(my$east, my$north), r=window*1000, npoints=100)
    nztm <- data.matrix(circle[,c("x", "y")])
    nzgd <- proj4::project(xy=nztm, proj=proj4string, inverse=TRUE)
    circle$long <- nzgd[,1]
    circle$lat <- nzgd[,2]

    # save selected farms for plot
    if (n >= 1) {
      farms <- rbind(farms, tibble(east=data_window$east, north=data_window$north,
                                 long=data_window$long, lat=data_window$lat,
                                 pasture=data_window$pasture_eaten,
                                 dist=data_window$dist, window=window, radius=as.factor(code)))
    }
    
    circles <- rbind(circles, tibble(long=circle$long, lat=circle$lat, radius=as.factor(code)))

    # save sample quantiles if enough data to be sensible
    if (n >= nmin) {
      
      # calculate quantile
      qr1 <- quantreg::rq(formula=pasture_eaten ~ 1, tau=0.9, data=data_window) # linear quantile regression
      se_method <- "boot" # how condience intervals are calculated, some methods more robust
      yqr1<- predict(qr1, tibble(east=my$east, north=my$north), interval="confidence", level=0.95, se=se_method)
      q90 <- quantile(data_window$pasture_eaten, 0.9, type=1) # also calc simple q90
  
      cat(file=stderr(), paste("yqr1 =", yqr1), "\n")
      # cat(file=stderr(), paste("q90 =", q90), "\n") # should be the same

      quants <- quantile(data_window$pasture_eaten, probs=probs, type=8) # see documentation for type=?
      sampcdf <- rbind(sampcdf, tibble(probs=probs, quants=quants, radius=as.factor(code)))
      samppdf <- rbind(samppdf, tibble(pasture=data_window$pasture_eaten, window=window,
                                       radius=as.factor(code),
                                       q=q90, qr=yqr1[1], qrlower=yqr1[2], qrupper=yqr1[3]))
      
    } # if n >= nmin

    # add a blank line (causes warnings but prevents errors)
    samppdf <- rbind(samppdf, tibble(pasture=NA, window=window,
                                 radius=as.factor(code),
                                 q=NA, qr=NA, qrlower=NA, qrupper=NA))

  } # next window size

  # biggest circle
  circle <- circle_fun(centre=c(my$east, my$north), r=max(windows)*1000, npoints=100)
  nztm <- data.matrix(circle[,c("x", "y")])
  nzgd <- proj4::project(xy=nztm, proj=proj4string, inverse=TRUE)
  circle$long <- nzgd[,1]
  circle$lat <- nzgd[,2]

  # return results as function for testing
  #calc <- function() list(data=data, circles=circles, circle=circle, farms=farms, sampcdf=sampcdf, samppdf=samppdf)
  
  # return results in a list
  return(list(data=data, circles=circles, circle=circle, farms=farms, 
              sampcdf=sampcdf, samppdf=samppdf))
  
}) # end reaction to elev chaning, calculation of calc <- list(results)

```

This chart shows the distribution of pasture and crop eaten within a 20km, 40km and 60km distance from your location. Potential is estimated as the 90th percentile – the level that only one out of ten farmers achieved more than. Around that level, the uncertainty band is shown as a shaded area. Charts won’t appear unless there are at least four farms in that group. The nitrogen adjustment, if selected, is calculated at a response of 10 kg DM per kg N fertiliser applied.

```{r output_plot, eval=TRUE, echo=FALSE}

#### create histograms ####

output$stacked_histogram <- renderPlot({

  samppdf <- calc()$samppdf # get data for histogram when calc() changes

  isolate({
    
    cat(file=stderr(), paste("render stacked histograms"), "\n")
    
    title_string <- paste("Pasture and Crop Eaten Near", my$name)

    # create empty plot
    stacked_histogram <- ggplot() +
      labs(title=title_string, y="Number of Farms\n", 
           x="Pasture and Crop Eaten, tonnes DM per ha", colour="Radius (km)") +
      ggthemes::theme_economist_white(base_size=12, horizontal=FALSE) +
      theme(axis.text.x=element_text(size=16, colour="grey35"),
            axis.text.y=element_text(size=16, colour="grey35"),
            strip.text=element_text(size=16, face="bold", colour="grey35"),
            plot.title=element_text(hjust=0.5, colour="grey35"),
            axis.title=element_text(size=16, face="bold", colour="grey35"),
            axis.title.x=element_text(size=18, face="bold", colour="grey35")
            ) +
      cowplot::panel_border(colour="grey35") +  
      theme(legend.position="none")
    
    # add histograms to empty plot
    if (nrow(drop_na(samppdf))>0) { 
      
      breaks <- my$breaks  

      # cat(file=stderr(), paste("xlim =", min(breaks), max(breaks)), "\n")
      
      stacked_histogram <- stacked_histogram +
        geom_rect(data=samppdf, mapping=aes(xmin=qrlower, xmax=qrupper, ymin=0, ymax=Inf), fill="lightcyan") +
        geom_histogram(data=samppdf, mapping=aes(x=pasture, colour=radius), fill=NA, size=1.1, binwidth=1) +
        geom_vline(data=samppdf, mapping=aes(xintercept=qr), size=1.5, colour="lightcyan4", alpha=0.2) +
        geom_vline(data=samppdf, mapping=aes(xintercept=q), size=1.5, colour="blue4") +
        geom_text(data=samppdf, mapping=aes(x=q, y=4, label=sprintf("%.1f t", q)),
                  colour="blue4", size=6, hjust=0, nudge_x=0.2) +
        facet_grid(radius ~ ., as.table=FALSE) + # as.table=FALSE reverses the order
        theme(strip.background=element_blank(), strip.text.y=element_text(angle=0)) +
        scale_x_continuous(breaks=breaks) +
        scale_y_continuous(breaks=NULL) +
        coord_cartesian(xlim=c(min(breaks),max(breaks)))
  
    } # end add histograms to empty plot
    
  }) # end isolate
  
  stacked_histogram

}) # end renderPlot
  
# show histogram
shinyUI(fluidPage(

    fluidRow(
      
      plotOutput("stacked_histogram")
      
    ) # end fluidRow
  
)) # end fluidPage

```

<!-- this comment was important!!!! -->
